---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
StepsPerDay <- aggregate(steps ~ date,activity,sum)
print(paste("Mean: ", mean(StepsPerDay$steps)))
print(paste("Median steps per day: ",median(StepsPerDay$steps)))
hist(StepsPerDay$steps,breaks = 20,xlab="Steps Per Day",ylab="Days", main = "")
```

## What is the average daily activity pattern?
```{r}
StepsPerInterval <- aggregate(steps ~ interval, activity,mean)
plot(StepsPerInterval$interval,StepsPerInterval$steps, xlab="5 Minute Intervals",ylab="Average Steps",type="l",main = "Average Steps Taken Throughout Days by 5 min Interval")

 maxInterval <- StepsPerInterval[StepsPerInterval$steps %in% max(StepsPerInterval$steps),]
 print(paste("Interval with the highest step count: ", maxInterval$interval))
```
## Imputing missing values

**Number of NAs in step data:**
```{r}
print(paste("Number of NAs in step data: ",sum(is.na(activity$steps))))
```

**Replace NA values in Step data with average steps for that interval:**
```{r}
library(dplyr)
noNa <- activity %>%  
  group_by(interval)%>% 
  mutate(steps=ifelse(is.na(steps),
                        as.integer(mean(steps,na.rm=TRUE)),
                        as.integer(steps)))

newStepsPerDay <- aggregate(steps ~ date,noNa,sum)
print(paste("New Mean: ", mean(newStepsPerDay$steps)))
print(paste("OLD Mean: ", mean(StepsPerDay$steps)))
print(paste("New Median: ",median(newStepsPerDay$steps)))
print(paste("OLD Median: ",median(StepsPerDay$steps)))
```
**The new mean is higher while the median steps per day is lower**  
**Overall it is evident there is A LOT more data!**
```{r}
par(mfrow = c(1, 2)) 
hist(StepsPerDay$steps,breaks = 20,xlab="Steps Per Day",ylab="Days", main = "With NAs", ylim = c(0,20))
hist(newStepsPerDay$steps,breaks = 20,xlab="Steps Per Day",ylab="Days", main = "NAs Replaced", ylim = c(0,20))
  
```

## Are there differences in activity patterns between weekdays and weekends?
